---
title: "Beetle Baby Maker- LABEL GENERATOR"
author: Timothy Gilbert
updated: 01/06/2022
output: flexdashboard::flex_dashboard
runtime: shiny_prerendered
#------------------------------------------------------------------------------------
#TO RUN, Click 'Run Document' at top of .Rmd document --> 'Generate Labels' --> 'Create Labels'
#IF APP SAYS CAN'T FIND FILE, TRY CLICKING ON 'Session'>'Set Working Directory'>'To Source File Location',
#THEN 'Run Document' AGAIN... Contact tsgilbert@arizona.edu for any help or feedback with application.

#This Rmarkdown document will read uploaded .csv beetle tracking sheet (designed by K.Matthiesen, D14) and create 
#beetle locality labels needed for carabid pinning. File column names in '2021 Carabid and IB Tracking.csv' 
#must not be changed [`Vial ID` and`# of individuals`]. Only works for Domain 14 - SRER & JORN as of 10/17/21 
#and will make a max of 30 localtity labels per trap location and a max of 250 locality labels in total.
#As of now does not make labels for carabids with unknown trap location (E,S,W)
#------------------------------------------------------------------------------------
---

```{r setup, include=FALSE}
# global env
## setting working dir() to this source file
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(shiny)
library(readr)
library(tidyverse)
library(data.table)
library(knitr)
library(kableExtra)
library(DT)
library(rmarkdown)
library(xfun)
#library(dplyr)

# update, without prompts for permission/clarification
#update.packages(ask = FALSE)


```

```{r data, include=FALSE}
# site localities data for display before making labels
data_f<- read_csv('data/neon_site_data.csv')

```

Inputs {.sidebar}
-------------------------------------------------------------

```{r}
# user interface inputs
fileInput("file1", label = "Upload '.csv' Tracking Sheet",accept = '.csv',placeholder = 'Carabid_Tracking.csv')

actionButton("button", "Generate Labels")

```

-------------------------------------------------------------


Upload .csv beetle tracking file, then **Generate Labels** FIRST to check for errors in both tabs. Then choose a font *size* and *style* and Click **'Create Labels'**. Labels will automatically open and save to working directory. Font size and style can be unique to each tab.

Note: Spacing for beetle locality labels designed for 4.5 size font and Arial style in current version with default spacing to NEON specifications.


-------------------------------------------------------------


```{r}

# display output window
htmlOutput("message")
```


Column {.tabset .tabset-fade}
-------------------------------------------------------------
### Beetle Locality Labels

```{r}
sliderInput("size", label = "Font Size Adjustment:",
            min = 2, max = 7, value = 4.5, step = 0.5)
sliderInput("size4", label = "Spacing Adjustment:",
            min = 10, max = 30, value = 20, step = 1)
selectInput("font", label = "Font Style Adjustment:",
            choices = c('Arial','Calibri','Cambria','Franklin Gothic',
                        'Futura','Garamond','Helvetica','Rockwell',
                        'Times New Roman','Verdana'))
actionButton("button2", "Create Locality Labels ", icon("file-export"))

# display output window
htmlOutput("labels")
```

### Beetle ID Labels
```{r}
# display output window
sliderInput("size2", label = "Font Size Adjustment:",
            min = 2, max = 7, value = 4.5, step = 0.5)
sliderInput("size3", label = "Spacing Adjustment:",
            min = 10, max = 22, value = 16, step = 1)
selectInput("font2", label = "Font Style Adjustment:",
            choices = c('Arial','Calibri','Cambria','Franklin Gothic',
                        'Futura','Garamond','Helvetica','Rockwell',
                        'Times New Roman','Verdana'))

actionButton("button3", "Create Identification Labels ", icon("file-export"))

htmlOutput("labels2")
```

```{r, context="server", echo=FALSE}
# reactive display based on uploaded file
data <- reactive({
  
    file <- input$file1
    ext <- tools::file_ext(file$datapath)

    req(file)
    validate(need(ext == "csv", "Please upload a csv file..."))

    chpe<- read_csv(file$datapath)
    
    site_data<- data_f %>% 
      select(elevation, decimalLatitude, decimalLongitude, sampleID)
    
    site_data<- site_data %>% 
      filter(!is.na(sampleID))

    sites_local<- unique(site_data)
    
chpe_f<- chpe %>% 
  filter(!is.na(`# of individuals`))
})

output$labels <- renderTable({
  req(file)
  
  chpe_f<- data()
  chpe_f %>% 
    select(`Vial ID`,`# of individuals`,Identification,`...9`) %>% 
    rename(Identifier = `...9`)
})

output$labels2 <- renderTable({
  req(file)
  
  chpe_f<- data()
  chpe_f %>% 
    select(`Vial ID`,`# of individuals`,Identification,`...9`) %>% 
    rename(Identifier = `...9`)
})

observeEvent(input$button,{
  
  chpe_f<- data()
  ## when button pushed... save uploaded file as .csv to working directory, then run code
  ## for making labels...
  write_csv(chpe_f,paste0('data/Carabid_and_IB_Tracking_',Sys.Date(),'.csv'),col_names = T)
  


  output$labels <- renderTable({
    
  withProgress(message = 'Processing Data...', value = 0, {
  # Increment the progress bar, and update the detail text.
  incProgress(.25, detail = paste("Processing...25%"))
    ## reading in beetle tracking sheet...
    chpe<- read_csv(paste0('data/Carabid_and_IB_Tracking_',Sys.Date(),'.csv'))
    chpe_f<- chpe %>% 
      filter(!is.na(`# of individuals`))
    ## check sheet for site... and dates...
    f_site<- substr(chpe_f$`Vial ID`[1],1,4)
    s_data<- data_f %>% 
      select(elevation, decimalLatitude, decimalLongitude, sampleID)
    s2_data<- s_data %>% 
      filter(!is.na(sampleID))
    site_data<- s2_data %>% 
      mutate(decimalLongitude = abs(decimalLongitude)) %>% 
      mutate(sampleID = substr(sampleID,1,10))
    
    sites_local<- unique(site_data)
    
    i=1
while (i < length(sites_local$sampleID)+1) {
  if (substr(sites_local$sampleID[i],10,10) == 'E'){ ##moving coordinate East
  ##-.000212 to Longitude (Westing)
  sites_local$decimalLongitude[i]<- (sites_local$decimalLongitude[i]-.000212)
  }
  if (substr(sites_local$sampleID[i],10,10) == 'W'){ ##moving coordinate West
  ##+.000212 to Longitude (Westing)
  sites_local$decimalLongitude[i]<- (sites_local$decimalLongitude[i]+.000212)
  }
  if (substr(sites_local$sampleID[i],10,10) == 'S'){ ##moving coordinate South
  ##-.000180 to Latitude (Northing))
  sites_local$decimalLatitude[i]<- (sites_local$decimalLatitude[i]-.000180)
  }
  i=i+1
 ## end of while statement 
}

## empty Table to fill up w/ data labels
local_labels <- data.table(BET_label_QC_check=character())

i=1 ## variable to move from each vial to another in loop
while (i < length(chpe_f$`Vial ID`)+1) { ## goes through every vial ID in excel file 'chpe_f'
  x=1
  ## looking at each location per sample
  while (x < (length(sites_local$sampleID))+1) {
    ## if vial id is equal to specific trap site local (SRER_005.W) will paste info in table
    if (substr(chpe_f$`Vial ID`[i],1,10) == substr(sites_local$sampleID[x],1,10)){
      elv<- sites_local$elevation[x]
      lat<- sites_local$decimalLatitude[x]
      long<- sites_local$decimalLongitude[x]
      ## formatting date for label
      date<- substr(chpe_f$`Vial ID`[i],12,19)
      date2<- as.Date(date, "%Y%m%d")
      f_date<- format(date2,"%d%h%Y")
      
      ## for Santa Rita Experimental Range
      if (substr(chpe_f$`Vial ID`[i],1,4) == 'SRER'){
        local_labels[[i,1]]<- paste0('USA, ARIZONA Pima Co.',
                                     '\n','Santa Rita Expm Rng. ',
                                     elv,'m ','\n','N',lat,' 路 W',long,
                                     '\n','Pitfall trap. ',f_date,'. AKaram.',
                                     '\n',substr(chpe_f$`Vial ID`[i],1,19))
      }
      ## for Jornada Experimental Range
      if (substr(chpe_f$`Vial ID`[i],1,4) == 'JORN'){
        local_labels[[i,1]]<- paste0('USA, NEW MEXICO Dona Ana Co.',
                                     '\n','Jornada Expm Rng. ',
                                     elv,'m ','\n','N',lat,' 路 W',long,
                                     '\n','Pitfall trap. ',f_date,'. AKaram.',
                                     '\n',substr(chpe_f$`Vial ID`[i],1,19))
      }
    }
    ## moves on to check another site local (x)
    x=x+1
  }
  i=i+1
}

## max of 30 labels per vial id
v=2
chpe_n<- list()
# number of rows dependent on `# of individuals` max
while (v < max(chpe_f$`# of individuals`)+1) {
  chpe_n[[v]]<- chpe_f %>% 
    filter(`# of individuals` > (v-1))
  v=v+1
}


#substr(sites_local$sampleID[x],1,10)
v=2
z=1
while (v < length(chpe_n)+1) {
  z=1 ## reset z to start vial id flow over
  while (z < length(chpe_n[[v]]$`Vial ID`)+1) { ## goes through every vial ID in excel file 'chpe_f'
    x=1
    ## looking at each location per sample
    while (x < (length(sites_local$sampleID))+1) {
      ## if vial id is equal to specific site local (SRER_005.W) will paste info in table- checks every local (x)
      if (substr(chpe_n[[v]]$`Vial ID`[z],1,10) == substr(sites_local$sampleID[x],1,10)){
        elv<- sites_local$elevation[x]
        lat<- sites_local$decimalLatitude[x]
        long2<- sites_local$decimalLongitude[x]
        long<- abs(long2)
        ## formatting date for label
        date<- substr(chpe_n[[v]]$`Vial ID`[z],12,19)
        date2<- as.Date(date, "%Y%m%d")
        f_date<- format(date2,"%d%h%Y")
       
        if (substr(chpe_n[[v]]$`Vial ID`[z],1,4) == 'SRER'){
          local_labels[[i,1]]<- paste0('USA, ARIZONA Pima Co.',
                                       '\n','Santa Rita Expm Rng. ',
                                       elv,'m ','\n','N',lat,' 路 W',long,
                                       '\n','Pitfall trap. ',f_date,'. AKaram.',
                                       '\n',substr(chpe_n[[v]]$`Vial ID`[z],1,19))
        }
        
        if (substr(chpe_n[[v]]$`Vial ID`[z],1,4) == 'JORN'){
          local_labels[[i,1]]<- paste0('USA, NEW MEXICO Dona Ana Co.',
                                       '\n','Jornada Expm Rng. ',
                                       elv,'m ','\n','N',lat,' 路 W',long,
                                       '\n','Pitfall trap. ',f_date,'. AKaram.',
                                       '\n',substr(chpe_n[[v]]$`Vial ID`[z],1,19))
        }
      }
      ## moves on to check another site local (x)
      x=x+1
    }
    
    i=i+1
    z=z+1
  }
  v=v+1
}

  incProgress(1, detail = paste("Done!"))
  }
  )

local_labels<- local_labels %>% 
  arrange(BET_label_QC_check)
})
  
  output$labels2 <- renderTable({
    ## reading in beetle tracking sheet...
    chpe<- read_csv(paste0('data/Carabid_and_IB_Tracking_',Sys.Date(),'.csv'))
    chpe_f<- chpe %>% 
      filter(!is.na(`# of individuals`))
    
    chpe_f<- chpe_f %>% 
    select(`Vial ID`,`# of individuals`,Identification,`...9`) %>% 
    rename(Identifier = `...9`)

## empty Table to fill up w/ data labels
local_labels2 <- data.table(BET_labels=character())

i=1 ## variable to move from each vial to another in loop
while (i < length(chpe_f$`Vial ID`)+1) { ## goes through every vial ID in excel file 'chpe_f'
  
  if (substr(chpe_f$`Vial ID`[i],12,13) == '20'){
    year<- substr(chpe_f$`Vial ID`[i],12,15)
  }
  if(substr(chpe_f$`Vial ID`[i],12,13) != '20'){
    year<- substr(chpe_f$`Vial ID`[i],10,13)
  }
  
  local_labels2[[i,1]]<- paste0('*',chpe_f$Identification[i],'*',' \n ','det. ',chpe_f$Identifier[i],' \n ',year)
  i=i+1
  }

## max of 30 labels per vial id
v=2
chpe_n<- list()
# number of rows dependent on `# of individuals` max
while (v < max(chpe_f$`# of individuals`)+1) {
  chpe_n[[v]]<- chpe_f %>% 
    filter(`# of individuals` > (v-1))
  v=v+1
}

v=2
z=1
while (v < length(chpe_n)+1) {
  z=1 ## reset z to start vial id flow over
  while (z < length(chpe_n[[v]]$`Vial ID`)+1) { ## goes through every vial ID in excel file 'chpe_f'
    
    if (substr(chpe_n[[v]]$`Vial ID`[z],12,13) == '20'){
      year2<- substr(chpe_n[[v]]$`Vial ID`[z],12,15)
    }
    ## plot location not known
    if (substr(chpe_n[[v]]$`Vial ID`[z],12,13) != '20'){
      year2<- substr(chpe_n[[v]]$`Vial ID`[z],10,13)
    }
        local_labels2[[i,1]]<- paste0('*',chpe_n[[v]]$Identification[z],'*',
                                      ' \n ','det. ',chpe_n[[v]]$Identifier[z],
                                      ' \n ',year2)
        i=i+1
        z=z+1
        }
  v=v+1
}

local_labels2<- local_labels2 %>% 
  arrange(BET_labels)
    })
})

## code to create locality labels
observeEvent(input$button2,{
  print('Creating Beetle Locality Labels, Processing...')

  size<- input$size
  font<- input$font
  size4<- input$size4
  Sys.sleep(.5)

  output$message <- renderText({
  print('Beetle Locality Labels Created, Opening pdf...')
  })
  
  # rendering labels with uploaded tracking sheet...
  rmarkdown::render("beetle_baby_maker.Rmd", output_format = "html_document")
  pagedown::chrome_print("beetle_baby_maker.html", output = "created_labels/BET_Locality_Labels.pdf")
  system2("open","created_labels/BET_Locality_Labels.pdf", wait = F)

  #system2("open","beetle_baby_maker.html")
})

## code to create identification labels
observeEvent(input$button3,{
  print('Creating Beetle Identifier Labels, Processing...')

  size2<- input$size2
  font2<- input$font2
  size3<- input$size3
  
  output$message <- renderText({
  print('Beetle Identifier Labels Created, Opening pdf...')
  })
  
  # rendering labels with uploaded tracking sheet...
  rmarkdown::render("beetle_identifier_labels.Rmd", output_format = "html_document")
  pagedown::chrome_print("beetle_identifier_labels.html",output = "created_labels/BET_ID_Labels.pdf")
  system2("open","created_labels/BET_ID_Labels.pdf")

})

```